A tool set for random number generation on GPUs in R(?)

### 2021-11-05 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
random exponential finished
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2947 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host using the MRG31k3p RNG. If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory. \pkg{gpuRandom} is able to create streams both on the host and on the GPU device. \fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior. Substreams are not created in \pkg{gpuRandom} because currently we use one stream per work-item. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# creating streams on CPU
library("gpuR")
library("gpuRandom")
myStreamsCpu <- createStreamsCpu(n=4, seed=12345)
t(myStreamsCpu)
##               [,1]       [,2]       [,3]       [,4]
## current.g1.1 12345  336690377  502033783  739421137
## current.g1.2 12345  597094797 1322587635 1475938232
## current.g1.3 12345 1245771585 1964121530  730262207
## current.g2.1 12345   85196284 1949818481 1630192198
## current.g2.2 12345  523477687 1607232546  324551134
## current.g2.3 12345 2094976052 1462898381  795289868
## initial.g1.1 12345  336690377  502033783  739421137
## initial.g1.2 12345  597094797 1322587635 1475938232
## initial.g1.3 12345 1245771585 1964121530  730262207
## initial.g2.1 12345   85196284 1949818481 1630192198
## initial.g2.2 12345  523477687 1607232546  324551134
## initial.g2.3 12345 2094976052 1462898381  795289868
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{createStreamsCpu} has two arguments:
\begin{itemize}
\itemsep0em 
  \item \code{n}: number of streams to create.
  \item \code{seed}: vector of length 6, recycled if shorter. Default is set to a vector of six ``12345''.
\end{itemize}
\fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device. We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams using \fct{createStreamsGpu}. Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# creating streams on GPU
myStreamsGpu1 = vclMatrix(myStreamsCpu)
myStreamsGpu2 = createStreamsGpu(n=4, seed=12345)
range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
## [1] 0 0
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{createStreamsGpu} also has the two arguments \code{n} and \code{seed}, \code{seed} is by default a vector of six ``12345''. \fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}. By setting the RNG status with the \proglang{R} function \fct{set.seed}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs. Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we make the sequence of streams repeatable.

@@

We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{gpuRandom::rnorm}. Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $512 \times 128$ work-items is around 120 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size goes larger.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streams <- createStreamsGpu(n = 512 * 128)
system.time(gpuRandom::rnorm(c(10000,10000), streams=streams, 
                             Nglobal=c(512,128), type="double"))
##    user  system elapsed 
##   0.138   0.000   0.137
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
system.time(matrix(stats::rnorm(10000^2),10000,10000))
##    user  system elapsed 
##   6.608   0.796   7.399
\end{verbatim}
\end{kframe}
\end{knitrout}

TO BE PUT IN APPENDIX OR DELETE? \\
In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}. The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution. We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streams <- createStreamsGpu(n =512*128)
normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
                                Nglobal=c(512,128), type="float")
avector<-as.vector(as.matrix(normal_matrix))
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
hist(avector,breaks=40, main = "histogram")
gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
\end{verbatim}
\end{kframe}
\end{knitrout}
\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
\end{figure}

@@

The test statistic calculated for each random table is $-\sum_{i,j}\log(n_{ij}!), i=(1,\dots, I), j=(1,\dots,J)$, (i.e., minus log-factorial of table), where $I$ and $J$ are the row and column number of the observed table. This test statistic can also be independently calculated for a table by \fct{gpuRandom::logfactSum}. Given an observed table and a number of replicates $B$, the Monte Carlo simulation for Fisher's exact test does the following steps: 
\begin{enumerate}
\setlength\itemsep{-0.1em}
  \item Calculate the test statistic for the observed table.
	\item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table, compute and sometimes save the test statistic from the random table. 
	\item Count the number of iterations (\emph{Counts}) that have test statistics less or equal to the one from the observed table.
	\item Estimate p-value using $\frac{1 + \emph{Counts}}{B + 1}$.
\end{enumerate}

Step 1 and step 2-3 are done on a GPU with two kernels enqueued sequentially. For step 2, \fct{gpuRandom::fisher.sim} adapted the function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random two-way tables with given marginal totals. The algorithm \citep[see][]{patefield1981efficient} samples the entries row by row, one at a time, conditional on the values of the entries already sampled. The conditional probabilities for the possible values of the next entry are updated dynamically each time a new entry is sampled. Then this entry is sampled by standard inversion of the cumulative distribution function, using one $U(0,1)$ random number.  For an $I \times J$ table, this requires $(I-1)(J-1)$ random numbers (the last column and last row do not need to be sampled). Finally, one can compute the test statistic for this newly sampled table. On a GPU, this step can be replicated say $n$ times in parallel by creating $n$ distinct random streams and launching $n$ separate work-items. Each work-item takes a random stream as input, performs all of Step 2, and returns the value of the test statistic and count on the GPU. Computing the p-value on the CPU (Step 4) is then a trivial operation.  Step 2 of saving test statistics from the random tables is made optional, which can reduce the run-time. %There are a lot of other more recent methods for sampling (larger) contingency tables, many of them use Markov Chain Monte Carlo. The use of streams and GPU would be quite different in that case.  See for example

We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU. The \proglang{R} outputs for testing on computer 2 is shown in the following.

% \begin{table}[H]
% \begin{tabular}{ |c|c|c|c|c|c| } 
% \hline
%            & host   &                     device \\ \hline
%  Computer1 & INtel Xenon W-2145 3.7 Ghz & Amd Radeon VII  \\ \hline
%  Computer2 & VCPU Intel Xenon Skylake 2.5ghz & VGPU Nvidia Tesla V100 \\ \hline
% \end{tabular}
% \caption{ Device information for run-time compare.\label{tab:device}}
% \end{table}

The 2-way contingency table \ref{tab:month} and table \ref{tab:week} consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}. %The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday. The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.

@@

month_gpu<-vclMatrix(month,type="integer")
result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
                   type="double", returnStatistics=TRUE,  Nglobal = c(256,128))
result_month$simNum
## [1] 1015808
result_month$counts
## [1] 410644
result_month$p.value
## [1] 0.4042541
\end{verbatim}
\end{kframe}
\end{knitrout}
We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.  
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
system.time(fisher.sim(month_gpu, 1e6, streams=streams, 
            type="double", returnStatistics=FALSE, Nglobal = c(256,128)))
##    user  system elapsed 
##   0.708   0.011   0.718
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1
week_GPU<-gpuR::vclMatrix(week,type="integer")
result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = c(256,128))
result_weekgpu$simNum
## [1] 10027008
result_weekgpu$counts
## [1] 1242
result_weekgpu$p.value
## [1] 0.0001239652
\end{verbatim}
\end{kframe}
\end{knitrout}
With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
            type="double", returnStatistics=FALSE,Nglobal = c(256,128)))
##    user  system elapsed 
##   4.350   0.017   4.364
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU.

@@

\end{knitrout}
% \begin{table}[H]
% \centering
% \begin{tabular}{ |c|c|c|c|c|c| } 
%  \hline
%  data &   B  & Intel CPU &  AMD GPU & NVIDIA GPU \\ \hline
%  Month & 1e6 & 0.403804 & 0.403507  & 0.403507     \\ \hline
%  Week &  1e7  & 0.0001251 & 0.0001274 & 0.0001274\\ \hline
% \end{tabular}
% \caption{ Summary of p-values obtained on different devices.\label{tab:summary}}
% \end{table}
% 
% 
% \begin{table}[H]
% \centering
% \begin{tabular}{ |c|c|c|c|c|c| } 
%  \hline
%  data &   B  & Intel CPU & sencha CPU & AMD GPU & NVIDIA GPU \\ \hline
%  Month & 1e6 & 9.74 & 14.58 & 2.28 & 0.31\\ \hline
%  Week &  1e7  & 60.19 & 90.38 & 10.82 & 1.90\\ \hline
% \end{tabular}
% \caption{ Run-time compare of Fisher's test simulation on different devices, all numbers are in seconds.\label{tab:compare time}}
% \end{table}

Figure \ref{fig4} visualizes the approximate sampling distributions of the test statistics from the two examples "Month" and "Week", by plotting a histogram for each of the example. The values of observed statistics are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot. 
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
\includegraphics[width=0.45\textwidth]{figure/fighistMonth-1} 
\includegraphics[width=0.45\textwidth]{figure/fighistMonth-2} 
\end{knitrout}
\caption{ Frequency distribution for test statistics from month and week data tables.\label{fig4}}
\end{figure}

@@

\begin{appendix}

\end{appendix}

>>>>

The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host using the MRG31k3p RNG. If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory. \pkg{gpuRandom} is able to create streams both on the host and on the GPU device. \fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior. Substreams are not created in \pkg{gpuRandom} because currently we use one stream per work-item. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# creating streams on CPU
library("gpuR")
library("gpuRandom")
myStreamsCpu <- createStreamsCpu(n=4, initial=12345)
t(myStreamsCpu)
##               [,1]       [,2]       [,3]       [,4]
## current.g1.1 12345  336690377  502033783  739421137
## current.g1.2 12345  597094797 1322587635 1475938232
## current.g1.3 12345 1245771585 1964121530  730262207
## current.g2.1 12345   85196284 1949818481 1630192198
## current.g2.2 12345  523477687 1607232546  324551134
## current.g2.3 12345 2094976052 1462898381  795289868
## initial.g1.1 12345  336690377  502033783  739421137
## initial.g1.2 12345  597094797 1322587635 1475938232
## initial.g1.3 12345 1245771585 1964121530  730262207
## initial.g2.1 12345   85196284 1949818481 1630192198
## initial.g2.2 12345  523477687 1607232546  324551134
## initial.g2.3 12345 2094976052 1462898381  795289868
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{createStreamsCpu} has two arguments:
\begin{itemize}
\itemsep0em 
  \item \code{n}: number of streams to create.
  \item \code{seed}: vector of length 6, recycled if shorter. Default is set to a vector of six ``12345''.
\end{itemize}
\fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device. We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams using \fct{createStreamsGpu}. Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# creating streams on GPU
myStreamsGpu1 = vclMatrix(myStreamsCpu)
myStreamsGpu2 = createStreamsGpu(n=4, initial=12345)
range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
## [1] 0 0
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{createStreamsGpu} also has the two arguments \code{n} and \code{seed}, \code{seed} is by default a vector of six ``12345''. \fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}. By setting the RNG status with the \proglang{R} function \fct{set.seed}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs. Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we make the sequence of streams repeatable.

@@

We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{gpuRandom::rnorm}. Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $512 \times 128$ work-items is around 120 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size goes larger.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streams <- createStreamsGpu(n = 512 * 128)
system.time(gpuRandom::rnorm(c(10000,10000), streams=streams, 
                             Nglobal=c(512,128), type="double"))
##    user  system elapsed 
##   0.133   0.000   0.133
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
system.time(matrix(stats::rnorm(10000^2),10000,10000))
##    user  system elapsed 
##   6.608   0.796   7.399
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

The test statistic calculated for each random table is $-\sum_{i,j}\log(n_{ij}!), i=(1,\dots, I), j=(1,\dots,J)$, (i.e., minus log-factorial of table), where $I$ and $J$ are the row and column number of the observed table. This test statistic can also be independently calculated for a table by \fct{gpuRandom::logfactSum}. Given an observed table and a number of replicates $B$, the Monte Carlo simulation for Fisher's exact test does the following steps: 
\begin{enumerate}
\setlength\itemsep{0em}
  \item Calculate the test statistic for the observed table.
	\item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table, compute and sometimes save the test statistic from the random table. 
	\item Count the number of iterations (\emph{Counts}) that have test statistics less or equal to the one from the observed table.
	\item Estimate p-value using $\frac{1 + \emph{Counts}}{B + 1}$.
\end{enumerate}

Step 1 and step 2-3 are done on a GPU with two kernels enqueued sequentially. For step 2, \fct{gpuRandom::fisher.sim} adapted the function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random two-way tables with given marginal totals. The algorithm \citep[see][]{patefield1981efficient} samples the entries row by row, one at a time, conditional on the values of the entries already sampled. The conditional probabilities for the possible values of the next entry are updated dynamically each time a new entry is sampled. Then this entry is sampled by standard inversion of the cumulative distribution function, using one $U(0,1)$ random number.  For an $I \times J$ table, this requires $(I-1)(J-1)$ random numbers (the last column and last row do not need to be sampled). Finally, one can compute the test statistic for this newly sampled table. On a GPU, this step can be replicated say $n$ times in parallel by creating $n$ distinct random streams and launching $n$ separate work-items. Each work-item takes a random stream as input, performs all of Step 2 and 3, and returns the value of the test statistic and count on the GPU. Computing the p-value on the CPU (Step 4) is then a trivial operation.  Step 2 of saving test statistics from the random tables is made optional, which can reduce the run-time. %There are a lot of other more recent methods for sampling (larger) contingency tables, many of them use Markov Chain Monte Carlo. The use of streams and GPU would be quite different in that case.  See for example

The arguments of the \fct{gpuRandom::fisher.sim} are described as follows:
\begin{itemize}
\itemsep0em 
  \item \code{x}: a vclMatrix.
  \item \code{N}: number of simulation runs.
  \item \code{streams}: streams.
  \item \code{type}: ``double'' or ``float'' or returned test statistics.
  \item \code{returnStatistics}: if TRUE, return all test statistics. 
  \item \code{Nglobal}: global index space. Default is set as $(64,16)$.
\end{itemize}
We show the advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU. The \proglang{R} outputs for testing on computer 2 is shown in the following.

The 2-way contingency tables %\ref{tab:month} and table \ref{tab:week} 
consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.

@@

month_gpu<-vclMatrix(month,type="integer")
result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
                   type="double", returnStatistics=TRUE,  Nglobal = c(256,128))
result_month$simNum
## [1] 1015808
result_month$counts
## [1] 410533
result_month$p.value
## [1] 0.4041449
\end{verbatim}
\end{kframe}
\end{knitrout}
We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.  
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
system.time(fisher.sim(month_gpu, 1e6, streams=streams, 
            type="double", returnStatistics=FALSE, Nglobal = c(256,128)))
##    user  system elapsed 
##   0.702   0.005   0.707
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1
week_GPU<-gpuR::vclMatrix(week,type="integer")
result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = c(256,128))
result_weekgpu$simNum
## [1] 10027008
result_weekgpu$counts
## [1] 1259
result_weekgpu$p.value
## [1] 0.0001256606
\end{verbatim}
\end{kframe}
\end{knitrout}
With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
            type="double", returnStatistics=FALSE,Nglobal = c(256,128)))
##    user  system elapsed 
##   4.324   0.011   4.332
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU.

@@

\end{knitrout}

\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
\includegraphics[width=0.45\textwidth]{figure/fighistMonth-1} 
\includegraphics[width=0.45\textwidth]{figure/fighistMonth-2} 
\end{knitrout}
\caption{Frequency distribution for test statistics from month and week data tables. Figure \ref{fig4} visualizes the approximate sampling distributions of the test statistics from the two examples "Month" and "Week", by plotting a histogram for each of the example. The values of observed statistics are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot. \label{fig4}}
\end{figure}

@@

\begin{appendix}
TO BE PUT IN APPENDIX OR DELETE? \\
In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}. The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution. We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streams <- createStreamsGpu(n =512*128)
normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
                                Nglobal=c(512,128), type="float")
avector<-as.vector(as.matrix(normal_matrix))
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
hist(avector,breaks=40, main = "histogram")
gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
\end{verbatim}
\end{kframe}
\end{knitrout}
\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
\end{figure}

The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday. The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]

\caption{\label{tab:monthdata}Month\label{tab:month}}
\centering
\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
\hline
  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
\hline
Jan & 29 & 55 & 172 & 46 & 39 & 73 & 48 & 183 & 77 & 103 & 102 & 174\\
\hline
Feb & 25 & 45 & 175 & 35 & 31 & 55 & 34 & 142 & 81 & 115 & 100 & 180\\
\hline
Mar & 31 & 48 & 182 & 41 & 47 & 72 & 40 & 200 & 86 & 90 & 96 & 180\\
\hline
Apr & 34 & 45 & 186 & 36 & 32 & 75 & 42 & 173 & 56 & 87 & 90 & 193\\
\hline
May & 33 & 40 & 187 & 46 & 24 & 80 & 35 & 180 & 75 & 91 & 100 & 197\\
\hline
Jun & 34 & 48 & 189 & 35 & 33 & 75 & 45 & 154 & 74 & 102 & 100 & 182\\
\hline
Jul & 26 & 43 & 198 & 34 & 21 & 74 & 36 & 179 & 79 & 86 & 92 & 193\\
\hline
Aug & 24 & 41 & 189 & 44 & 43 & 62 & 48 & 183 & 88 & 109 & 94 & 194\\
\hline
Sept & 34 & 44 & 147 & 40 & 37 & 66 & 36 & 158 & 73 & 112 & 103 & 196\\
\hline
Oct & 25 & 43 & 207 & 45 & 31 & 65 & 49 & 181 & 77 & 108 & 115 & 220\\
\hline
Nov & 36 & 55 & 188 & 39 & 39 & 62 & 43 & 144 & 68 & 98 & 79 & 173\\
\hline
Dec & 23 & 48 & 196 & 31 & 31 & 71 & 31 & 177 & 86 & 86 & 73 & 156\\
\hline
\end{tabular}
\end{table}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]

\caption{\label{tab:weekdata}Week\label{tab:week}}
\centering
\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
\hline
  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
\hline
Mon & 30 & 34 & 173 & 37 & 23 & 80 & 49 & 191 & 83 & 122 & 109 & 216\\
\hline
Tue & 60 & 121 & 383 & 80 & 83 & 131 & 71 & 349 & 146 & 164 & 168 & 352\\
\hline
Wed & 51 & 106 & 417 & 92 & 73 & 145 & 72 & 333 & 136 & 179 & 196 & 351\\
\hline
Thu & 60 & 86 & 362 & 69 & 74 & 120 & 85 & 326 & 132 & 220 & 187 & 359\\
\hline
Fri & 52 & 94 & 347 & 87 & 59 & 123 & 68 & 323 & 145 & 170 & 166 & 345\\
\hline
Sat & 52 & 63 & 323 & 67 & 64 & 135 & 73 & 316 & 170 & 189 & 188 & 357\\
\hline
Sun & 49 & 51 & 211 & 40 & 32 & 96 & 69 & 216 & 108 & 143 & 130 & 258\\
\hline
\end{tabular}
\end{table}

\end{knitrout}
\end{appendix}
